### 2020-10-19 tex/initial-presentation.tex
Complete initial presentation.

\documentclass[aspectratio=169]{beamer}

\usetheme{metropolis}

\usefonttheme{professionalfonts}
\usepackage[familydefault,light]{Chivo}

\input{common.tex}

\date{\formatdate{20}{10}{2020}}

\institute{supervised by Prof. Dr. Ruth Breu}

\begin{document}
  \maketitle

  \begin{frame}{Outline}
    \begin{itemize}
      \item Motivation
      \item Low-Code Architecture
      \item Model-Driven Architecture
      \item Criticisms
      \item Planned Project Procedure
    \end{itemize}
  \end{frame}

  \begin{frame}{Motivation}
    \begin{itemize}
      \item What can and can't low-code and model-driven tools do?
      \item Are they a viable alternative to traditional development?
      \item When to choose one approach over the other?
    \end{itemize}
  \end{frame}

  \begin{frame}{Low-Code Architecture}
    \begin{itemize}
      \item provides pre-built application components
      \item graphical user interface for creating both the
            application logic as well as the user interface
      \item typically aimed at end-users rather than developers
    \end{itemize}
  \end{frame}

  \begin{frame}{Model-Driven Architecture}
    \begin{itemize}
      \item provides a set of guidelines for the structuring of specifications
      \item code (fully or partially) generated from models, e.g. from UML diagrams
      \item aimed at developers to speed up development
    \end{itemize}
  \end{frame}

  \begin{frame}{Criticisms}
    \begin{itemize}
      \item Low-Code Architecture
        \begin{itemize}
          \item Unsuitable for implementing scalable and mission-critical applications.
          \item Increase in unsupported applications built by shadow IT,
                i.e. applications which are not controlled by a company's IT department.
        \end{itemize}
      \item Model-Driven Architecture
        \begin{itemize}
          \item UML diagrams lack details included in the code itself.
          \item “the Code is the design” - Should models be derived from code instead of code from models?
        \end{itemize}
      \item Do these approaches actually make development easier and cheaper?
    \end{itemize}
  \end{frame}

  \begin{frame}{Planned Project Procedure}
    \begin{itemize}
      \item Evaluate a set of Low-Code tools.
      \item Evaluate a set of Model-Driven tools.
      \item For each tool, determine what it can and can't do.
      \item Compare the two sets with each other and draw some conclusions.
    \end{itemize}
  \end{frame}
\end{document}