### 2021-01-02 tex/paper.tex
Extend paper.

\documentclass[runningheads]{llncs}

\linespread{1.1}

\raggedbottom

\usepackage{float}

\input{common.tex}

\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{nameref}

\addto\extrasenglish{
  \def\sectionautorefname{Section}
}

\usepackage[
  style   = numeric,
  sorting = none,
]{biblatex}
\bibliography{paper}

\begin{document}

\institute{University of Innsbruck, Austria}

\maketitle

\vspace{8em}

\begin{abstract}

In this paper we present a survey conducted by comparing low-code tools with model-driven tools for developing software in order to get a better understanding of the advantages and drawbacks of both approaches and whether they are a viable alternative to traditional software development.

\keywords{Low-Code \and Model-Driven \and Development}

\end{abstract}

\newpage

\section{Introduction}
\label{sec:introduction}

In this paper, we compare Low-Code Platforms to Model-Driven Architecture and try to determine when and where to use them over Model-Driven Architecture or vice-versa.

In order to make clear to the reader the difference between the two approaches we firstly provide definitions and characteristics of each paradigm in the following sections.

\subsection{What is Model-Driven Architecture?}
\label{ssec:what_is_model_driven}

With Model-Driven Architecture or Model-Driven Engineering, the goal is to standardise on models in a given domain. This is done in order to reduce code duplication and speed up development. Especially when bootstrapping a new project, it is helpful keep boilerplate code to a minimum. In addition to reducing development time, Model-Driven Engineering also tries to increase productivity by maximising compatibility between systems by using the aforementioned standardised models.~\cite{wiki:model_driven_engineering} Generally, Model-Driven Architectures are geared towards developers, i.e. people who also have a good understanding of the underlying programming language(s) of the respective architecture. One common way Model-Driven Engineering is used in practice is code generation. Using code generation can speed up many processes: For example, given an API specification (e.g. using the industry standard OpenAPI~\cite{openapis}), one can generate all code that is necessary for a client for this API. Swagger~\cite{swagger} is one such tool which can generate API clients for many different programming languages. There are also tools which can generate a complete class hierarchy from an UML diagram; many IDEs offer such functionality either natively or via plug-ins. These are only a few examples of what falls into the category of Model-Driven Architecture, the main point is that all of them help developers reach their goal faster than doing the same tasks manually.

\subsection{What is a Low-Code Platform?}
\label{ssec:what_is_a_low_code_platform}

A low-code platform, as the name implies, requires a minimal amount of code to be written. Similarly to conventional programming (i.e. using a programming language), an IDE is used for creating programs using low-code platforms. Often, the IDE is part of the low-code platform itself instead of a standalone program as is the case for conventional programming. In most cases, the IDE of the platform and the low-code platform cannot be used separately so the term low-code platform is also commonly used to refer to the IDE interface itself. Low-code platforms are targeted towards users with little to no programming experience, which means the actual programming of a low-code system is done mainly via visual building blocks. These building blocks contain pre-built templates and functionality and can be combined to form a coherent user interface. Oftentimes, custom functionality can be implemented using a click-and-drag pattern between building blocks and by assigning an action accordingly. The available building blocks vary by platform which makes them less portable than a system written using a programming language. This is one common point of critique on low-code platforms.

Another sub-category of low-code platforms are so called no-code platforms~\cite{wiki:no_code_development_platform}. Where to draw the line between low-code and no-code is not clearly defined, especially since low-code platforms can sometimes be used in a way that cannot be differentiated from no-code platforms. For the purposes of this paper, low-code platforms are general purpose visual programming languages~\cite{wiki:visual_programming_language} aimed at both developer and non-developer users while no-code platforms are closer to software as a service (SaaS) platforms targeted mostly (or only) towards non-developers. Another way to look at it is that low-code platforms still give developers the freedom to add custom building blocks for their specific use case and no-code solutions are usually purpose-made for a specific line of business or a specific type of department. Some examples include monday.com~\cite{wiki:monday_com}, a no-code platform for project management tasks; FileMaker~\cite{wiki:filemaker}, which provides a database engine and a corresponding no-code platform for creating graphical user interfaces for interacting with databases; and Airtable~\cite{wiki:airtable}, a no-code platform for creating databases connected to interactive spreadsheets.

\section{Evaluation of Low-Code Tools}
\label{sec:evaluation_of_low_code_tools}

\subsection{Selection of Low-Code Tools}
\label{subsec:choosing_low_code_tools}

Before evaluating low-code tools, we had to first determine which tools to evaluate. Given the ever evolving landscape of low-code platforms, there are countless companies offering low-code solutions. The following is an incomplete list of low-code platforms we found:

\begin{itemize}
  \item Oracle APEX (Application Express) (originally released as Oracle Flows in 2000) by Oracle (founded in 1977)
  \item OutSystems (founded in 2001)
  \item Mendix (founded in 2005), a subsidiary of Siemens
  \item Simplifier (founded in 2012 as iTiZZiMO, renamed 2019 to Simplifier, the same name as its low-code platform)
  \item OSBP (Open Standard Buiness Platform), an open-source low-code platform (developed since 2016) under the umbrella of the Eclipse Foundation
  \item Corteza Low Code, part of the open-source Corteza Project (released in 2019), originally developed by Crust Technology
\end{itemize}

We wanted our evaluation to contain at least one open-source platform (e.g. OSBP), one platform backed by a well-known corporation (e.g. Oracle), one platform by a relatively new/unknown corporation (e.g. Simplifier), one relatively old/well-established platform (e.g. Oracle APEX or OutSystems) and one relatively new platform (e.g. Simplifier or Corteza). With the given list of low-code platforms, we could satisfy this requirement.

\section{Evaluation of Model-Driven Tools}
\label{sec:evaluation_of_model_driven_tools}

\section{Findings}
\label{sec:findings}

\section{Conclusion}
\label{sec:conclusion}

\newpage
\printbibliography

\end{document}